---
title: "Covariates data preparation dogs Tana"
author: "Katerina"
format: html
editor: visual
---

```{r}
library(RStoolbox)    
library(tidyverse)
library(sf)         
library(randomForest)   
library(terra)         
library(exactextractr) 
library(units)
library(pls)
library(raster)
library(emmeans)
library(factoextra)

```

# Supervised classification of CUA satellite image

```{r}
#load satelite image
tif <- brick("CUA_ras.tif") %>% setMinMax()

# Load training polygons and align CRS to the raster
train <- st_read("training_cls4.shp", quiet = TRUE) %>%
  mutate(class = as.factor(class)) %>%
  st_transform(crs = sf::st_crs(raster::crs(tif)))


#supervised classification
sc <- superClass(tif, trainData = train, responseCol = "class", model = "rf") 

#check it out and save in file
plot(sc$map)
sc$classMapping
writeRaster(sc$map, "classified_CUA.tif", overwrite = TRUE) #file not saved in github, too large

plot(tif)
```

# Calculate proportions of urban classes per fokontany

```{r}
# load files; classified raster layer, polygons counted
layer_cls <- raster::raster(sc$map) #or we can load classified tif, if sc is no more in the environment 

fkt_counted <- st_read("Poly_fkt_counted_lar.shp") %>%
  st_transform(crs = sf::st_crs(raster::crs(layer_cls))) %>%
  arrange(tolower(Name))

# superpose the layer and polygons and compute class-wise coverage directly
coverage_fkt_counted <- exact_extract(
  layer_cls,
  fkt_counted,
  function(values, coverage_fraction, ...) {
    tapply(coverage_fraction, values, sum, na.rm = TRUE)
  },
  progress = FALSE)

coverage_fkt_counted

coverage_fkt_counted <- t(coverage_fkt_counted) #flipping the matrix

#if i want to approximate real area in meters (but CAVE, there is a boundary error, it gives slightly higher values than QGIS)
coverage_fkt_counted <- coverage_fkt_counted * (prod(res(layer_cls)))

colnames(coverage_fkt_counted) <- sc$classMapping$class

#arrange it into a neat dataframe
urban_str_fkt <- as.data.frame(coverage_fkt_counted) %>%
  mutate(
    fkt = fkt_counted$Name,
    id_fkt = 1:19,
    sum = water + green + industrial + urban,
    p_water = water/sum,
    p_green = green/sum,
    p_industrial = industrial / sum,
    p_urban = urban/sum,
    sum_check = p_water + p_green + p_industrial + p_urban)
View(urban_str_fkt)

write.csv(urban_str_fkt, file = "urban_str_fkt.csv")
```

# Calculate proportions of urban classes per patch

```{r}
transect_counted <- st_read("Transect_counted.shp")%>% st_transform(crs = crs(layer_cls)) %>% arrange(tolower(Name))

# superpose the layer and the polygons and count per polygon coverage 
coverage_transect_counted <- exact_extract(
  layer_cls,
  transect_counted,
  function(values, coverage_fraction, ...) {
    values <- factor(values, levels = sc$classMapping$classID)
    out <- tapply(coverage_fraction, values, sum, na.rm = TRUE)
    out[is.na(out)] <- 0
    out
  },
  progress = FALSE)

coverage_transect_counted <- t(coverage_transect_counted)
colnames(coverage_transect_counted) <- sc$classMapping$class

urban_str_transect <- as.data.frame(coverage_transect_counted) %>%
  mutate(
    name_transect = transect_counted$Name,
    id_transect = 1:102,
    sum = water + green + industrial + urban,
    p_water = water/sum,
    p_green = green/sum,
    p_industrial = industrial / sum,
    p_urban = urban/sum,
    sum_check = p_water + p_green + p_industrial + p_urban)
View(urban_str_transect)

write.csv(urban_str_transect, file = "urban_str_transect.csv")

```

# Calculate proportions of urban classes in all fokontany

This is to see what the proportions of different urban classes are doing. Please note that the sampled fokontany are not a subset of the total 192 fokontany; they are QGIS-created polygons surrounding the paths counted. Sometimes they are almost equal to administrative polygons, other times they are very different. The purpose of creating sampled fokontany polygons was to stay true to the real area that was counted (that is why, in the sitecovs_fkt, we have "area_admin" and area_real".)

```{r}
#load shp of all fokontany of the CUA (obtained from OCHA)
fkt_all <- st_read("CUA_FKT.shp") %>% st_transform(crs(layer_cls)) %>% arrange(tolower(ADM4_EN))

# superpose the layer and the polygons and count per polygon coverage 
coverage_fkt_all <- exact_extract(
  layer_cls,
  fkt_all,
  function(values, coverage_fraction, ...) {
    values <- factor(values, levels = sc$classMapping$classID)
    out <- tapply(coverage_fraction, values, sum, na.rm = TRUE)
    out[is.na(out)] <- 0
    out
  },
  progress = FALSE)

coverage_fkt_all <- t(coverage_fkt_all)
colnames(coverage_fkt_all) <- sc$classMapping$class

urban_str_fkt_all <- as.data.frame(coverage_fkt_all) %>%
  mutate(
    fkt = fkt_all$ADM4_EN,
    area = fkt_all$area1,
    sum = water + green + industrial + urban,
    p_water = water/sum,
    p_green = green/sum,
    p_industrial = industrial / sum,
    p_urban = urban/sum,
    sum_check = p_water + p_green + p_urban + p_industrial)

View(urban_str_fkt_all)

write.csv(urban_str_fkt_all, file = "urban_str_all_fkt.csv")

fkt_all
```

# Create PCA-based urban indexes

Based on proportions of urban classes across the whole CUA, we calculate the principal component analysis to introduce index 1 and 2, corresponding to PC1 and 2, thus decreasing the number of future covariates for the model and avoiding collinearity between variables. It turns out that PC1 explains 60% of the variability, which is nice.

```{r}

# Run PCA 
pca_result <- prcomp(urban_str_fkt_all[, c("p_green", "p_urban", "p_industrial", "p_water")], center = TRUE, scale. = TRUE)

# Explained variance; first two components
var_explained <- (pca_result$sdev^2) / sum(pca_result$sdev^2) * 100
var_explained[1:2]

scores <- predict(pca_result, newdata = urban_str_fkt_all[, c("p_green","p_urban","p_industrial","p_water")])

#we create "raw" form and scaled form of the index, for different purposes downstream
urban_str_fkt_all <- urban_str_fkt_all %>% 
  mutate(
    urban_index1 = scores[,1],
    urban_index2 = scores[,2],
    index1_s = scale(urban_index1),
    index2_s = scale(urban_index2))

# Save raw PCA scores for all fokontany
write.csv(urban_str_fkt_all, file = "urban_str_all_fkt.csv", row.names = FALSE)

View(urban_str_fkt_all)


```

# Compile all covariates available per fokontany

```{r}
# load demographic data acquired from QGIS and in the field
demography_fkt <- read_csv("Demography_fkt_counted.csv")

# use the PCA1 and PCA2 scores above to calculate urban indexes specific for the areas counted 
scores_fkt_counted <- predict(pca_result, newdata = urban_str_fkt[, c("p_green","p_urban","p_industrial","p_water")])

# Attach counted-site PCA indexes directly to the counted urban table
urban_str_fkt <- urban_str_fkt %>%
  mutate(
    index1_raw = as.numeric(scores_fkt_counted[,1]),
    index2_raw = as.numeric(scores_fkt_counted[,2]),
    index1 = as.numeric(scale(scores_fkt_counted[,1])),
    index2 = as.numeric(scale(scores_fkt_counted[,2]))
  )

sitecovs_fkt <- left_join(demography_fkt, urban_str_fkt, by = "fkt") %>%
  mutate(
         pop_dens_s = scale(population / area_admin), #adding human population density
         area_s = scale(area_real),
         pcc_s = scale(pcc_total),
         length_s = scale(length))

View(sitecovs_fkt)
write.csv(sitecovs_fkt, file = "sitecovs_fkt.csv")
```

# Compile all covariates available per path

note that some variables are only known at fokontany level, so transects are nested per fokontany in some covariates

```{r}
char_transect <- read_csv("Transect_counted.csv") %>% left_join(urban_str_transect, by = "id_transect")

scores_transect_counted <- predict(pca_result, newdata = urban_str_transect[, c("p_green","p_urban","p_industrial","p_water")])

sitecovs_transect <- left_join(char_transect, demography_fkt, by = "fkt") %>%
  mutate(index1 = scale(scores_transect_counted[,1]),
         index2 = scale(scores_transect_counted[,2]),
         pop_dens_s = scale(population / area_admin), #adding human population density
         area_s = scale(area_real),
         pcc_s = scale(pcc_total),
         length_s = scale(length))

View(sitecovs_transect)
write.csv(sitecovs_transect, file = "sitecovs_transect.csv")
```

# Spatial data exploration; dogs per urban class

```{r}
dogs <- st_read("Dogs5days.shp") %>% st_transform(crs(layer_cls))

# Both packages are loaded so we must be explicit
layer_cls_terra <- terra::rast(layer_cls)

# extract raster class value at each dog GPS point
ex <- terra::extract(layer_cls_terra, terra::vect(dogs), ID = FALSE)  

# attach class values to points 
dogs$class <- ex[,1]

# count by class
dogs_per_zone <- dogs %>%
  st_set_geometry(NULL) %>%     # drop geometry to make a data frame
  group_by(class) %>%
  summarise(n = n()) %>%
  arrange(class) %>%
  mutate(class = as.factor(recode(class,
    `1` = "green",
    `2` = "industrial",
    `3` = "urban", 
    `4` = "water")))

dogs_per_zone$area_zone <- as.numeric(c(sum(sitecovs_fkt$urban), sum(sitecovs_fkt$industrial),sum(sitecovs_fkt$green), sum(sitecovs_fkt$water), "NA"))

dogs_per_zone$density_zone <- (dogs_per_zone$n) / (dogs_per_zone$area_zone) *2000 #per ha but divided by 5 (because all 5 repetitions are included)

print(dogs_per_zone)


#test of zone effect

m <- glm(n ~ class + offset(log(area_zone)), family = poisson, data = dogs_per_zone)

anova(m, test = "Chisq")

summary(m)

em <- emm <- emmeans(m, "class", type = "response")
summary(emm)  
pairs(emm, type = "response", adjust = "fdr")

plot(emm, comparisons = TRUE, horizontal = TRUE,
     main = "Estimated dog rate per unit area (95% CI)")

```
