---
title: "Covariates data preparation dogs Tana"
author: "Katerina"
format: html
editor: visual
---

```{r}
library(RStoolbox)    
library(tidyverse)
library(sf)         
library(randomForest)   
library(terra)         
library(exactextractr) 
library(units)
library(pls)
library(raster)
library(emmeans)

```

# Supervised classification of CUA satellite image

```{r}
#load satelite image
tif <- brick("CUA_ras.tif") %>% setMinMax()

# Load training polygons and align CRS to the raster
train <- st_read("training_cls.shp", quiet = TRUE) %>%
  mutate(class = as.factor(class)) %>%
  st_transform(crs = sf::st_crs(raster::crs(tif)))


#supervised classification
sc <- superClass(tif, trainData = train, responseCol = "class", model = "rf") 

#check it out and save in file
plot(sc$map)
sc$classMapping
writeRaster(sc$map, "classified3.tif", overwrite = TRUE)


```

# Calculate proportions of urban classes per fokontany

```{r}
# load files; classified raster layer, polygons counted
layer_cls <- raster::raster(sc$map) #or we can load classified tif, if sc is no more in the environment 

fkt_counted <- st_read("Poly_fkt_counted.shp") %>%
  st_transform(crs = sf::st_crs(raster::crs(layer_cls))) %>%
  arrange(tolower(Name))

# superpose the layer and polygons and compute class-wise coverage directly
coverage_fkt_counted <- exact_extract(
  layer_cls,
  fkt_counted,
  function(values, coverage_fraction, ...) {
    tapply(coverage_fraction, values, sum, na.rm = TRUE)
  },
  progress = FALSE)

coverage_fkt_counted

coverage_fkt_counted <- t(coverage_fkt_counted) #flipping the matrix

#if i want to approximate real area in meters (but cave, there is a boundary error, it gives slightly higher values than QGIS)
coverage_fkt_counted <- coverage_fkt_counted * (prod(res(layer_cls)))

colnames(coverage_fkt_counted) <- sc$classMapping$class

#arrange it into a neat dataframe
urban_str_fkt <- as.data.frame(coverage_fkt_counted) %>%
  mutate(
    fkt = fkt_counted$Name,
    id_fkt = 1:19,
    sum = rowSums(across(where(is.numeric)), na.rm = TRUE),
    p_water = water/sum,
    p_ricefield = ricefield/sum,
    p_industrial = industrial / sum,
    p_sparse = sparse_urb/sum,
    p_compact = compact_ur / sum,
    sum_check = p_water + p_ricefield + p_sparse + p_industrial + p_compact)
View(urban_str_fkt)

write.csv(urban_str_fkt, file = "urban_str_fkt.csv")
```

# Calculate proportions of urban classes per patch

```{r}
transect_counted <- st_read("Transect_counted.shp")%>% st_transform(crs = crs(layer_cls)) %>% arrange(tolower(Name))

# superpose the layer and the polygons and count per polygon coverage 
coverage_transect_counted <- exact_extract(
  layer_cls,
  transect_counted,
  function(values, coverage_fraction, ...) {
    tapply(coverage_fraction, values, sum, na.rm = TRUE)
  },
  progress = FALSE)

coverage_transect_counted <- t(coverage_transect_counted)
colnames(coverage_transect_counted) <- sc$classMapping$class

urban_str_transect <- as.data.frame(coverage_transect_counted) %>%
  mutate(
    name_transect = transect_counted$Name,
    id_transect = 1:102,
    sum = rowSums(across(where(is.numeric)), na.rm = TRUE),
    p_water = water/sum,
    p_ricefield = ricefield/sum,
    p_industrial = industrial / sum,
    p_sparse = sparse_urb/sum,
    p_compact = compact_ur / sum,
    sum_check = p_water + p_ricefield + p_sparse + p_industrial + p_compact)
View(urban_str_transect)

write.csv(urban_str_transect, file = "urban_str_transect.csv")

```

# Calculate proportions of urban classes in all fokontany

```{r}
fkt_all <- st_read("CUA_FKT.shp") %>% st_transform(crs(layer_cls)) %>% arrange(tolower(ADM4_EN))

# superpose the layer and the polygons and count per polygon coverage 
coverage_fkt_all <- exact_extract(
  layer_cls,
  fkt_all,
  function(values, coverage_fraction, ...) {
    tapply(coverage_fraction, values, sum, na.rm = TRUE)
  },
  progress = FALSE)

coverage_fkt_all <- t(coverage_fkt_all)
colnames(coverage_fkt_all) <- sc$classMapping$class

urban_str_fkt_all <- as.data.frame(coverage_fkt_all) %>%
  mutate(
    fkt = fkt_all$ADM4_EN,
    sum = rowSums(across(where(is.numeric)), na.rm = TRUE),
    p_water = water/sum,
    p_ricefield = ricefield/sum,
    p_industrial = industrial / sum,
    p_sparse = sparse_urb/sum,
    p_compact = compact_ur / sum,
    sum_check = p_water + p_ricefield + p_sparse + p_industrial + p_compact)

View(urban_str_fkt_all)

write.csv(urban_str_fkt_all, file = "urban_str_all_fkt.csv")
```

# Create PCA-based urban indexes

Based on proportions of urban classes across all city

```{r}

# Run PCA 
pca_result <- prcomp(urban_str_fkt_all[, c("p_compact", "p_sparse", "p_ricefield", "p_industrial", "p_water")], center = TRUE, scale. = TRUE)

# PCA scores
pca_result$x

# Explained variance; first two components
var_explained <- (pca_result$sdev^2) / sum(pca_result$sdev^2) * 100
var_explained[1:2]

scores <- predict(pca_result, newdata = urban_str_fkt_all[, c("p_compact","p_sparse","p_ricefield","p_industrial","p_water")])

urban_str_fkt_all <- urban_str_fkt_all %>% mutate(urban_index1 = scores[,1], urban_index2 = scores[,2])

# Save raw PCA scores for all fokontany
write.csv(urban_str_fkt_all, file = "urban_str_all_fkt.csv", row.names = FALSE)

View(urban_str_fkt_all)

```

# Compile all covariates available per fokontany

```{r}
# load demographic data acquired from QGIS and in the field
demography_fkt <- read_csv("Demography_fkt_counted.csv")

# use the PCA1 and PCA2 scores above to calculate urban indexes specific for the areas counted 
scores_fkt_counted <- predict(pca_result, newdata = urban_str_fkt[, c("p_compact","p_sparse","p_ricefield","p_industrial","p_water")])

# Attach counted-site PCA indexes directly to the counted urban table
urban_str_fkt <- urban_str_fkt %>%
  mutate(
    index1 = as.numeric(scale(scores_fkt_counted[,1])),
    index2 = as.numeric(scale(scores_fkt_counted[,2]))
  )

sitecovs_fkt <- left_join(demography_fkt, urban_str_fkt, by = "fkt") %>%
  mutate(
         pop_dens_s = scale(population / area_admin), #adding human population density
         area_s = scale(area_real),
         pcc_s = scale(pcc_total),
         length_s = scale(length))

View(sitecovs_fkt)
write.csv(sitecovs_fkt, file = "sitecovs_fkt.csv")
```

# Compile all covariates available per path

note that some variables are only known at fokontany level, so transects are nested per fokontany in some covariates

```{r}
char_transect <- read_csv("Transect_counted.csv") %>% left_join(urban_str_transect, by = "id_transect")

scores_transect_counted <- predict(pca_result, newdata = urban_str_transect[, c("p_compact","p_sparse","p_ricefield","p_industrial","p_water")])

sitecovs_transect <- left_join(char_transect, demography_fkt, by = "fkt") %>%
  mutate(index1 = scale(scores_transect_counted[,1]),
         index2 = scale(scores_transect_counted[,2]),
         pop_dens_s = scale(population / area_admin), #adding human population density
         area_s = scale(area_real),
         pcc_s = scale(pcc_total),
         length_s = scale(length))

View(sitecovs_transect)
write.csv(sitecovs_transect, file = "sitecovs_transect.csv")
```

# Spatial data exploration; dogs per urban class

```{r}
dogs <- st_read("Dogs5days.shp") %>% st_transform(crs(layer_cls))

# Both packages are loaded so we must be explicit
layer_cls_terra <- terra::rast(layer_cls)

# extract raster class value at each dog GPS point
ex <- terra::extract(layer_cls_terra, terra::vect(dogs), ID = FALSE)  

# attach class values to points 
dogs$class <- ex[,1]

# count by class
dogs_per_zone <- dogs %>%
  st_set_geometry(NULL) %>%     # drop geometry to make a data frame
  group_by(class) %>%
  summarise(n = n()) %>%
  arrange(class) %>%
  mutate(class = as.factor(recode(class,
    `1` = "compact_ur",
    `2` = "industrial",
    `3` = "ricefield", 
    `4` = "sparse_urb",
    `5`  = "water")))

dogs_per_zone$area_zone <- as.numeric(c(sum(sitecovs_transect$compact_ur), sum(sitecovs_transect$industrial),sum(sitecovs_transect$ricefield), sum(sitecovs_transect$sparse_urb), sum(sitecovs_transect$water), "NA"))

dogs_per_zone$density_zone <- (dogs_per_zone$n) / (dogs_per_zone$area_zone) *2000 #per ha but divided by 5 (because all 5 repetitions are included)

print(dogs_per_zone)


#test of zone effect

m <- glm(n ~ class + offset(log(area_zone)), family = poisson, data = dogs_per_zone)

anova(m, test = "Chisq")

summary(m)

em <- emm <- emmeans(m, "class", type = "response")
summary(emm)  
pairs(emm, type = "response", adjust = "fdr")

plot(emm, comparisons = TRUE, horizontal = TRUE,
     main = "Estimated dog rate per unit area (95% CI)")

```
