---
title: "INLA N-mixture"
author: "Katerina"
format: html
editor: visual
---

Packages

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library("INLA") 

```

## Models using fokontany as site

### Load Data; fokontany as site

The files we load here were created in the workflows "count data preparation" and "covariate data preparation". If you did not create the objects, they are provided in the github repository, so you can just load the csv filed required here.

Note that INLA gets too complicated when there is different number of surveys per site, so we use the first five days / visits at each site, not making use of 9 repetitions in one the the 19 sites.

```{r}
sitecovs_fkt <- read.csv("sitecovs_fkt.csv") %>% mutate(
  se_classe = recode(as.character(Seclasse),
                       "Riche" = 5,
                       "Moyen" = 3,
                       "Pauvre" = 1) %>% as.numeric()  %>% scale() %>% as.numeric(),)

View(sitecovs_fkt)

obs_matrix_fkt_5 <- as.matrix(read_csv("obs_matrix_fkt.csv", col_select = 2:6))

mean(obs_matrix_fkt_5)
range(obs_matrix_fkt_5)
```

#### Model 1

```{r}
counts_mdata1 <- inla.mdata(
  obs_matrix_fkt_5,
  1,                              
  sitecovs_fkt$length_s,               
  sitecovs_fkt$pop_dens_s,                  
  sitecovs_fkt$pcc_s)

data_list1 <- list(
  counts_mdata1 = counts_mdata1, 
  index1 = sitecovs_fkt$index1,
  index2  = sitecovs_fkt$index2)

model_1 <- inla(
  counts_mdata1 ~ 1 + index1 + index2,   
  data = data_list1,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_1)

print(model_1$ok)

inla.nmix.lambda.fitted(model_1)

inla.nmix.lambda.fitted(model_1)$fitted.summary %>% pull(mean.lambda) %>% sum() #sum per all sites
inla.nmix.lambda.fitted(model_1)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_1)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_1$summary.fixed["(Intercept)","mean"]) #average detection probability

```

#### Model 2

```{r}
counts_mdata2 <- inla.mdata(
  obs_matrix_fkt_5,
  1,                              
  sitecovs_fkt$length_s,               
  sitecovs_fkt$pop_dens_s,
  sitecovs_fkt$se_classe)

data_list2 <- list(
  counts_mdata2 = counts_mdata2, 
  index1 = sitecovs_fkt$index1,
  index2 = sitecovs_fkt$index2)

model_2 <- inla(
  counts_mdata2 ~ 1 + index1 +index2,   
  data = data_list2,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_2)

print(model_2$ok)

inla.nmix.lambda.fitted(model_2)

inla.nmix.lambda.fitted(model_2)$fitted.summary %>% pull(mean.lambda) %>% sum()
inla.nmix.lambda.fitted(model_2)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_2)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_2$summary.fixed["(Intercept)","mean"]) 
```

#### Model 3

```{r}
counts_mdata3 <- inla.mdata(
  obs_matrix_fkt_5,
  1,                              
  sitecovs_fkt$length_s,               
  sitecovs_fkt$pop_dens_s)

data_list3 <- list(
  counts_mdata3 = counts_mdata3, 
  index1 = sitecovs_fkt$index1)

model_3 <- inla(
  counts_mdata3 ~ 1 + index1,   
  data = data_list3,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_3)

print(model_3$ok)

inla.nmix.lambda.fitted(model_3)

inla.nmix.lambda.fitted(model_3)$fitted.summary %>% pull(mean.lambda) %>% sum() 
inla.nmix.lambda.fitted(model_3)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_3)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_3$summary.fixed["(Intercept)","mean"]) 
```

```{r}
results_m3 <- as.data.frame(inla.nmix.lambda.fitted(model_3))%>% mutate(
  Fokontany = sitecovs_fkt$fkt,
  area_ha = sitecovs_fkt$area_real / 10000,
  population = sitecovs_fkt$population,
  dogs_density = fitted.summary.mean.lambda / area_ha, 
  dog_per_km = fitted.summary.mean.lambda / sitecovs_fkt$length)

write.csv(results_m3, file = "resultsm3.csv")
```

#### Model 3 b

```{r}
# Separate candidate model: add a site-level iid effect.
# This is not replacing Model 3; it is fitted as an additional model for comparison.

sitecovs_fkt$site_id <- seq_len(nrow(sitecovs_fkt))

counts_mdata3_site <- inla.mdata(
  obs_matrix_fkt_5,
  1,
  sitecovs_fkt$length_s,
  sitecovs_fkt$pop_dens_s
)

data_list3_site <- list(
  counts_mdata3_site = counts_mdata3_site,
  index1 = sitecovs_fkt$index1,
  site_id = sitecovs_fkt$site_id
)

model_3_site_iid <- inla(
  counts_mdata3_site ~ 1 + index1 + 
    f(site_id, model = "iid",
      hyper = list(prec = list(prior = "pc.prec", param = c(1, 0.01)))),
  data = data_list3_site,
  family = "nmix",

  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),

  control.family = list(
    hyper = list(
      theta1 = list(prior = "loggamma", param = c(1, 0.5))
    )
  ),

  control.compute = list(cpo = TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor = TRUE, config = TRUE)
)

summary(model_3_site_iid)

print(model_3_site_iid$ok)


```

#### Model 4

```{r}
counts_mdata4 <- inla.mdata(
  obs_matrix_fkt_5,
  1,                              
  sitecovs_fkt$area_s,               
  scale(sitecovs_fkt$p_urban),
  scale(sitecovs_fkt$p_green))
  
data_list4 <- list(
  counts_mdata4 = counts_mdata4,
  index1 = sitecovs_fkt$index1)

model_4 <- inla(
  counts_mdata4 ~ 1 + index1,   
  data = data_list4,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_4)

print(model_4$ok)

inla.nmix.lambda.fitted(model_4)

inla.nmix.lambda.fitted(model_4)$fitted.summary %>% pull(mean.lambda) %>% sum()
inla.nmix.lambda.fitted(model_4)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_4)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_4$summary.fixed["(Intercept)","mean"])
```

```         
```

#### Model 5

```{r}
counts_mdata5 <- inla.mdata(
   obs_matrix_fkt_5,
  1,                              
  sitecovs_fkt$length_s,               
  scale(sitecovs_fkt$p_urban),
  scale(sitecovs_fkt$p_green))
  
data_list5 <- list(
  counts_mdata5 = counts_mdata5,
  index1 = sitecovs_fkt$index1)

model_5 <- inla(
  counts_mdata5 ~ 1 + index1,   
  data = data_list5,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_5)

print(model_5$ok)

inla.nmix.lambda.fitted(model_5)

inla.nmix.lambda.fitted(model_5)$fitted.summary %>% pull(mean.lambda) %>% sum() 
inla.nmix.lambda.fitted(model_5)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_5)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_5$summary.fixed["(Intercept)","mean"])
```

#### Model 6

```{r}
counts_mdata6 <- inla.mdata(
  obs_matrix_fkt_5,
  1,                              
  sitecovs_fkt$length_s,
  sitecovs_fkt$index1)

data_list6 <- list(
  counts_mdata6 = counts_mdata6)

model_6 <- inla(
  counts_mdata6 ~ 1,
  data = data_list6,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_6)

print(model_6$ok)

inla.nmix.lambda.fitted(model_6)

inla.nmix.lambda.fitted(model_6)$fitted.summary %>% pull(mean.lambda) %>% sum()
inla.nmix.lambda.fitted(model_6)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_6)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_6$summary.fixed["(Intercept)","mean"])
```

#### Model 7

```{r}
counts_mdata7 <- inla.mdata(
  obs_matrix_fkt_5,
  1,                              
  sitecovs_fkt$length_s,
  scale(sitecovs_fkt$p_urban),
  scale(sitecovs_fkt$p_green),
  scale(sitecovs_fkt$p_industrial))

data_list7 <- list(
  counts_mdata7 = counts_mdata7,
  index1 = sitecovs_fkt$index1)

model_7 <- inla(
  counts_mdata7 ~ 1 + index1,   
  data = data_list7,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 4.77, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_7)

print(model_7$ok)

inla.nmix.lambda.fitted(model_7)

inla.nmix.lambda.fitted(model_7)$fitted.summary %>% pull(mean.lambda) %>% sum()
inla.nmix.lambda.fitted(model_7)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_7)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_7$summary.fixed["(Intercept)","mean"])
```

## Models using transect as site

```{r}
sitecovs_transect <- read.csv("sitecovs_transect.csv")

obs_matrix_transect_5 <- as.matrix(read_csv("obs_matrix_transect.csv", col_select = 2:6))

View(sitecovs_transect)
obs_matrix_transect_5

mean(obs_matrix_transect_5)
range(obs_matrix_transect_5)

```

#### Model 10

```{r}
counts_mdata10 <- inla.mdata(
  obs_matrix_transect_5,
  1,                              
  sitecovs_transect$length_s,               
  sitecovs_transect$pop_dens_s)

data_list10 <- list(
  counts_mdata10 = counts_mdata10, 
  index1 = sitecovs_transect$index1,
  index2  = sitecovs_transect$index2)

model_10 <- inla(
  counts_mdata10 ~ 1 + index1 + index2,   
  data = data_list10,
  family = "nmixnb",
  
  control.fixed = list(mean.intercept = 3, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_10)

print(model_10$ok)

inla.nmix.lambda.fitted(model_10)

inla.nmix.lambda.fitted(model_10)$fitted.summary %>% pull(mean.lambda) %>% sum()
inla.nmix.lambda.fitted(model_10)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_10)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_10$summary.fixed["(Intercept)","mean"]) 
```

#### Model 11

```{r}
counts_mdata11 <- inla.mdata(
  obs_matrix_transect_5,
  1,                              
  sitecovs_transect$length_s,               
  sitecovs_transect$index1)

data_list11 <- list(
  counts_mdata11 = counts_mdata11, 
  index2 = sitecovs_transect$index2)

model_11 <- inla(
  counts_mdata11 ~ 1 + index2,   
  data = data_list11,
  family = "nmix",
  
  control.fixed = list(mean.intercept = 3, prec.intercept = 0.1),
  
  control.family = list(
    hyper = list(
      theta1 = list(prior="loggamma", param=c(1, 0.5)))),
  
  control.compute = list(cpo=TRUE, dic = TRUE, waic = TRUE, return.marginals.predictor=TRUE, config = TRUE))

summary(model_11)

print(model_11$ok)

inla.nmix.lambda.fitted(model_11)

inla.nmix.lambda.fitted(model_11)$fitted.summary %>% pull(mean.lambda) %>% sum()
inla.nmix.lambda.fitted(model_11)$fitted.summary %>% pull(quant025.lambda) %>% sum()
inla.nmix.lambda.fitted(model_11)$fitted.summary %>% pull(quant975.lambda) %>% sum()
plogis(model_11$summary.fixed["(Intercept)","mean"]) 
```

# Model Comparison

```{r}
# Helper function to extract metrics
extract_metrics <- function(model, model_name) {
  dic <- model$dic$dic
  waic <- model$waic$waic
  mlik <- model$mlik[1]
  plml <- mean(log(pmin(model$cpo$cpo, 1e10)), na.rm = TRUE) 
  
  return(data.frame(
    model = model_name,
    DIC = dic,
    WAIC = waic,
    MLIK = mlik,
    PLML = plml,
    stringsAsFactors = FALSE
  ))
}

comparison <- list(
  extract_metrics(model_1, "Model 1"),
  extract_metrics(model_2, "Model 2"),
  extract_metrics(model_3, "Model 3"),
  extract_metrics(model_4, "Model 4"),
  extract_metrics(model_5, "Model 5"),
  extract_metrics(model_10, "Model 6"),
  extract_metrics(model_11, "Model 7"))

dplyr::bind_rows(comparison)

```

Recall;

DIC, WAIC, PLML: Lower values indicate better fit

MLIK: Higher values indicate better model (marginal likelihood); better predictive capacity

## PPC for the best models

Models 2 and 3 have overall the best criteria and ecologcially makes sense;

positive influence of index1 on detection =\> where index1 is negative (=high proportion of compact urban structure) detection is low, where positive (= high proportion of ricefields), detection is higher.

```{r}
set.seed(123)

# Hierarchical PPC for Model 2
# Abundance uncertainty: lambda_i from inla.nmix.lambda.fitted()
# Detection uncertainty: p_i = logit^{-1}(theta0 + theta1 * index1_i)

lambda_sum <- inla.nmix.lambda.fitted(model_2)$fitted.summary #abundance summary
lambda_mean <- lambda_sum$mean.lambda # posterior mean abundance
lambda_sd <- pmax(lambda_sum$sd.lambda, 1e-8) #sd per site, slight positivity forced

idx1 <- as.numeric(data_list2$index1)

fixed_sum <- model_2$summary.fixed
theta0_mean <- fixed_sum["(Intercept)", "mean"] #posterior mean of the detection intercept (baseline detection)
theta0_sd <- pmax(fixed_sum["(Intercept)", "sd"], 1e-8) #ps sd det inter, forced slightly pos
theta1_mean <- fixed_sum["index1", "mean"] # posterior mean of effect of index1
theta1_sd <- pmax(fixed_sum["index1", "sd"], 1e-8) #ps sd index1

y_obs <- obs_matrix_fkt_5

sum(obs_matrix_fkt_5)

site_names <- as.character(sitecovs_fkt$fkt) 
n_sites <- nrow(y_obs)
n_visits <- ncol(y_obs)
n_samp <- 2000

# Convert each site's posterior lambda into lognormal parameters (forcing slightly positive values)
sdlog <- sqrt(log1p((lambda_sd^2) / pmax(lambda_mean^2, 1e-12)))
meanlog <- log(pmax(lambda_mean, 1e-12)) - 0.5 * sdlog^2

obs_total <- sum(y_obs)
obs_site_tot <- rowSums(y_obs) #each site's total across visits

# create storage objects for simulated results across iterations
rep_total <- numeric(n_samp)
rep_site_tot <- matrix(NA_real_, nrow = n_samp, ncol = n_sites) #empty matrix with 1 column per site and 1 row per each posterior sample


# Simulation loop
### each s is one replicated dataset

for (s in seq_len(n_samp)) {
  lambda_draw <- rlnorm(n_sites, meanlog = meanlog, sdlog = pmax(sdlog, 1e-8))

  # Normal approximation for fixed effects / detection 
  theta0_draw <- rnorm(1, mean = theta0_mean, sd = theta0_sd)
  theta1_draw <- rnorm(1, mean = theta1_mean, sd = theta1_sd)
  p_draw <- plogis(theta0_draw + theta1_draw * idx1)
  p_draw <- pmin(pmax(p_draw, 1e-8), 1 - 1e-8)

  N_draw <- rpois(n_sites, lambda_draw) #latent true abundance

  #replicate the fieldwork under the principles established for model2
  y_rep <- matrix(NA_real_, nrow = n_sites, ncol = n_visits)
  for (i in seq_len(n_sites)) {
    y_rep[i, ] <- rbinom(n_visits, size = N_draw[i], prob = p_draw[i])
  }

  rep_site_tot[s, ] <- rowSums(y_rep)
  rep_total[s] <- sum(y_rep)
}

# Overall count comparison (all 19 sites pooled)
overall_ppc <- tibble(
  observed_total = obs_total,
  predicted_median = median(rep_total),
  predicted_q025 = quantile(rep_total, 0.025),
  predicted_q975 = quantile(rep_total, 0.975),
  p_rep_ge_obs = mean(rep_total >= obs_total),
  p_rep_le_obs = mean(rep_total <= obs_total)
)
print(overall_ppc)

ggplot(tibble(total_rep = rep_total), aes(x = total_rep)) +
  geom_histogram(bins = 35, fill = "#1b6ca8", color = "white") +
  geom_vline(xintercept = obs_total, color = "#cc3d3d", linewidth = 1) +
  labs(
    title = "PPC for Overall Count Across 19 Sites (Model 2)",
    subtitle = "Blue: replicated totals, red line: observed total",
    x = "Replicated overall count",
    y = "Frequency"
  ) +
  theme_minimal()

# Site-level totals comparison
site_ppc <- tibble(
  site = site_names,
  observed_total = obs_site_tot,
  rep_median = apply(rep_site_tot, 2, median, na.rm = TRUE),
  rep_q025 = apply(rep_site_tot, 2, quantile, probs = 0.025, na.rm = TRUE),
  rep_q975 = apply(rep_site_tot, 2, quantile, probs = 0.975, na.rm = TRUE)
)

ggplot(site_ppc %>% mutate(site = factor(site, levels = site[order(observed_total)])), aes(x = site)) +
  geom_pointrange(aes(y = rep_median, ymin = rep_q025, ymax = rep_q975), color = "#1b6ca8") +
  geom_point(aes(y = observed_total), color = "#cc3d3d", size = 2) +
  coord_flip() +
  labs(
    title = "Site-level Totals PPC (Model 2)",
    subtitle = "Blue: replicated median and 95% interval, red: observed",
    x = "Site",
    y = "Total count across visits"
  ) +
  theme_minimal()
```

```{r}
set.seed(123)

# Hierarchical PPC for Model 3
# Abundance uncertainty: lambda_i from inla.nmix.lambda.fitted()
# Detection uncertainty: p_i = logit^{-1}(theta0 + theta1 * index1_i)

lambda_sum <- inla.nmix.lambda.fitted(model_3)$fitted.summary #abundance summary
lambda_mean <- lambda_sum$mean.lambda # posterior mean abundance
lambda_sd <- pmax(lambda_sum$sd.lambda, 1e-8) #sd per site, slight positivity forced

idx1 <- as.numeric(data_list2$index1)

fixed_sum <- model_2$summary.fixed
theta0_mean <- fixed_sum["(Intercept)", "mean"] #posterior mean of the detection intercept (baseline detection)
theta0_sd <- pmax(fixed_sum["(Intercept)", "sd"], 1e-8) #ps sd det inter, forced slightly pos
theta1_mean <- fixed_sum["index1", "mean"] # posterior mean of effect of index1
theta1_sd <- pmax(fixed_sum["index1", "sd"], 1e-8) #ps sd index1

y_obs <- obs_matrix_fkt_5

sum(obs_matrix_fkt_5)

site_names <- as.character(sitecovs_fkt$fkt) 
n_sites <- nrow(y_obs)
n_visits <- ncol(y_obs)
n_samp <- 2000

# Convert each site's posterior lambda into lognormal parameters (forcing slightly positive values)
sdlog <- sqrt(log1p((lambda_sd^2) / pmax(lambda_mean^2, 1e-12)))
meanlog <- log(pmax(lambda_mean, 1e-12)) - 0.5 * sdlog^2

obs_total <- sum(y_obs)
obs_site_tot <- rowSums(y_obs) #each site's total across visits

# create storage objects for simulated results across iterations
rep_total <- numeric(n_samp)
rep_site_tot <- matrix(NA_real_, nrow = n_samp, ncol = n_sites) #empty matrix with 1 column per site and 1 row per each posterior sample


# Simulation loop
### each s is one replicated dataset

for (s in seq_len(n_samp)) {
  lambda_draw <- rlnorm(n_sites, meanlog = meanlog, sdlog = pmax(sdlog, 1e-8))

  # Normal approximation for fixed effects / detection 
  theta0_draw <- rnorm(1, mean = theta0_mean, sd = theta0_sd)
  theta1_draw <- rnorm(1, mean = theta1_mean, sd = theta1_sd)
  p_draw <- plogis(theta0_draw + theta1_draw * idx1)
  p_draw <- pmin(pmax(p_draw, 1e-8), 1 - 1e-8)

  N_draw <- rpois(n_sites, lambda_draw) #latent true abundance

  #replicate the fieldwork under the principles established for model2
  y_rep <- matrix(NA_real_, nrow = n_sites, ncol = n_visits)
  for (i in seq_len(n_sites)) {
    y_rep[i, ] <- rbinom(n_visits, size = N_draw[i], prob = p_draw[i])
  }

  rep_site_tot[s, ] <- rowSums(y_rep)
  rep_total[s] <- sum(y_rep)
}

# Overall count comparison (all 19 sites pooled)
overall_ppc <- tibble(
  observed_total = obs_total,
  predicted_median = median(rep_total),
  predicted_q025 = quantile(rep_total, 0.025),
  predicted_q975 = quantile(rep_total, 0.975),
  p_rep_ge_obs = mean(rep_total >= obs_total),
  p_rep_le_obs = mean(rep_total <= obs_total)
)
print(overall_ppc)

ggplot(tibble(total_rep = rep_total), aes(x = total_rep)) +
  geom_histogram(bins = 35, fill = "#1b6ca8", color = "white") +
  geom_vline(xintercept = obs_total, color = "#cc3d3d", linewidth = 1) +
  labs(
    title = "PPC for Overall Count Across 19 Sites (Model 3)",
    subtitle = "Blue: replicated totals, red line: observed total",
    x = "Replicated overall count",
    y = "Frequency"
  ) +
  theme_minimal()

# Site-level totals comparison
site_ppc <- tibble(
  site = site_names,
  observed_total = obs_site_tot,
  rep_median = apply(rep_site_tot, 2, median, na.rm = TRUE),
  rep_q025 = apply(rep_site_tot, 2, quantile, probs = 0.025, na.rm = TRUE),
  rep_q975 = apply(rep_site_tot, 2, quantile, probs = 0.975, na.rm = TRUE)
)

ggplot(site_ppc %>% mutate(site = factor(site, levels = site[order(observed_total)])), aes(x = site)) +
  geom_pointrange(aes(y = rep_median, ymin = rep_q025, ymax = rep_q975), color = "#1b6ca8") +
  geom_point(aes(y = observed_total), color = "#cc3d3d", size = 2) +
  coord_flip() +
  labs(
    title = "Site-level Totals PPC (Model 3)",
    subtitle = "Blue: replicated median and 95% interval, red: observed",
    x = "Site",
    y = "Total count across visits"
  ) +
  theme_minimal()
```

## Graph Observed counts and estimated abundance per fokontany

```{r}
# Observed count distribution across 5 visits and estimated abundance (lambda) by fokontany
site_names <- as.character(sitecovs_fkt[["fkt"]])
site_lengths <- suppressWarnings(as.numeric(sitecovs_fkt[["length"]]))
site_order <- unique(site_names[!is.na(site_names) & site_names != "" & is.finite(site_lengths)][order(site_lengths[!is.na(site_names) & site_names != "" & is.finite(site_lengths)])])

lambda_plot_df <- inla.nmix.lambda.fitted(model_3)$fitted.summary %>%
  as_tibble() %>%
  mutate(
    site = factor(sitecovs_fkt$fkt, levels = site_order)
  )

obs_plot_df <- as_tibble(obs_matrix_fkt_5) %>%
  mutate(site = factor(sitecovs_fkt$fkt, levels = site_order)) %>%
  pivot_longer(
    cols = -site,
    names_to = "visit",
    values_to = "count"
  ) %>%
  mutate(visit = factor(visit, levels = paste0("...", 1:5), labels = paste("Visit", 1:5))) %>%
  filter(!is.na(site), !is.na(count))

lambda_plot_df <- lambda_plot_df %>%
  filter(!is.na(site), is.finite(mean.lambda), is.finite(quant025.lambda), is.finite(quant975.lambda))

ggplot() +
  geom_point(
    data = obs_plot_df,
    aes(x = site, y = count),
    position = position_jitter(width = 0.06, height = 0),
    size = 1.1,
    color = "#0b3c5d",
    alpha = 0.85,
    na.rm = TRUE
  ) +
  geom_pointrange(
    data = lambda_plot_df,
    aes(
      x = site,
      y = mean.lambda,
      ymin = quant025.lambda,
      ymax = quant975.lambda
    ),
    color = "#cc3d3d",
    fatten = 1.2,
    linewidth = 0.5,
    position = position_nudge(x = 0.28),
    na.rm = TRUE
  ) +
  coord_flip() +
  labs(
    title = "Observed counts and estimated abundance by fokontany",
    x = "Fokontany",
    y = "Count / estimated abundance"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    legend.position = "bottom"
  )
```

## Graph observed counts versus predicted counts

```{r}
# Observed count dispersion (box-whisker) vs predicted observed counts (lambda * p)
# Note: CI for lambda*p is an approximation combining lambda and detection-coefficient quantiles.

coef_df <- model_3$summary.fixed %>%
  as.data.frame() %>%
  rownames_to_column("term")

beta0_mean <- coef_df %>% filter(term == "(Intercept)") %>% pull(mean)
beta1_mean <- coef_df %>% filter(term == "index1") %>% pull(mean)
beta2_mean <- coef_df %>% filter(term == "index2") %>% pull(mean)

beta0_lwr <- coef_df %>% filter(term == "(Intercept)") %>% pull(`0.025quant`)
beta1_lwr <- coef_df %>% filter(term == "index1") %>% pull(`0.025quant`)
beta2_lwr <- coef_df %>% filter(term == "index2") %>% pull(`0.025quant`)

beta0_upr <- coef_df %>% filter(term == "(Intercept)") %>% pull(`0.975quant`)
beta1_upr <- coef_df %>% filter(term == "index1") %>% pull(`0.975quant`)
beta2_upr <- coef_df %>% filter(term == "index2") %>% pull(`0.975quant`)

p_df <- tibble(
  site = factor(sitecovs_fkt$fkt, levels = site_order),
  index1 = as.numeric(sitecovs_fkt$index1)
) %>%
  mutate(
    p_mean = plogis(beta0_mean + beta1_mean * index1),
    p_lwr = plogis(beta0_lwr + beta1_lwr * index1),
    p_upr = plogis(beta0_upr + beta1_upr * index1)
  ) %>%
  filter(!is.na(site), is.finite(p_mean), is.finite(p_lwr), is.finite(p_upr))

pred_obs_df <- lambda_plot_df %>%
  left_join(p_df, by = "site") %>%
  transmute(
    site,
    mean_lp = mean.lambda * p_mean,
    lwr_lp = quant025.lambda * p_lwr,
    upr_lp = quant975.lambda * p_upr
  ) %>%
  filter(!is.na(site), is.finite(mean_lp), is.finite(lwr_lp), is.finite(upr_lp))

obs_box_df <- obs_plot_df %>%
  filter(!is.na(site), is.finite(count))

ggplot() +
  geom_boxplot(
    data = obs_box_df,
    aes(x = site, y = count, group = site),
    width = 0.55,
    fill = "#d9e6f2",
    color = "#4f6d7a",
    outlier.shape = NA,
    na.rm = TRUE
  ) +
  geom_pointrange(
    data = pred_obs_df,
    aes(x = site, y = mean_lp, ymin = lwr_lp, ymax = upr_lp),
    color = "#2e8b57",
    fatten = 1.2,
    linewidth = 0.5,
    position = position_nudge(x = 0.28),
    na.rm = TRUE
  ) +
  coord_flip() +
  labs(
    title = "Observed count dispersion vs predicted observed counts (Model 3)",
    x = "Fokontany",
    y = "Observed counts / predicted observed counts"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    legend.position = "bottom"
  )
```

# Extrapolation CUA

### Extrapolation based on urban proportions

```{r}

# sample totals
sample_area <- sum(sitecovs_fkt$area_real) / 10000
sample_dogs <- inla.nmix.lambda.fitted(model_3)$fitted.summary %>% pull(mean.lambda) %>% sum()
sample_density <- sample_dogs / sample_area

# distribution of classes in sample and their proportions
sample_green <- sum(sitecovs_fkt$area_real *sitecovs_fkt$p_green) / 10000
sample_industrial <- sum(sitecovs_fkt$area_real*sitecovs_fkt$p_industrial) / 10000
sample_urban <- sum(sitecovs_fkt$area_real*sitecovs_fkt$p_urban) / 10000
sample_water <- sum(sitecovs_fkt$area_real*sitecovs_fkt$p_water) / 10000

p_sample_green <- sample_green / sample_area
p_sample_industrial <- sample_industrial / sample_area
p_sample_urban <- sample_urban / sample_area
p_sample_water <- sample_water / sample_area

sampled_class_props <- data.frame(
  class = c("green","industrial","urban","water"),
  area_ha = c(sample_green, sample_industrial, sample_urban, sample_water),
  proportion = c(p_sample_green, p_sample_industrial, p_sample_urban, p_sample_water),
  emmeans_rate = c(172, 1618, 4767, 470)
) %>% mutate(
  dog_share = (proportion * emmeans_rate) / sum(proportion * emmeans_rate),
  dogs_in_class = dog_share * sample_dogs,
  density_per_class = dogs_in_class / area_ha
)

sampled_class_props

dogs_per_fkt <- urban_str_fkt_all %>% transmute(
  fkt = fkt,
  area_ha = area / 10000,
  dogs_green = 0.14 * area_ha * p_green, 
  dogs_industrial = 1.32 * area_ha * p_industrial,
  dogs_urban = 3.89 * area_ha * p_urban,
  dogs_water = 0.38 * area_ha * p_water,
  dogs_total = dogs_green + dogs_urban + dogs_industrial + dogs_water
)

sum(dogs_per_fkt$dogs_total)

```

### Propagation of uncertainty in the extrapolate

```{r}

set.seed(123)
n_sim <- 5000

urban_str_fkt_all <- read.csv("urban_str_all_fkt.csv")

# 1) Draw sampled abundance from model_3 posterior uncertainty
fit_sum <- inla.nmix.lambda.fitted(model_3)$fitted.summary
# expected columns: mean.lambda, sd.lambda

mu_sd_to_lnorm <- function(m, s) {
  # Convert mean/sd on natural scale -> lognormal parameters
  sigma2 <- log1p((s^2) / (m^2))
  mu <- log(m) - 0.5 * sigma2
  c(mu = mu, sdlog = sqrt(sigma2))
}

pars <- t(mapply(mu_sd_to_lnorm, fit_sum$mean.lambda, fit_sum$sd.lambda))

# lambda draws per site and simulation
lambda_draws <- sapply(seq_len(nrow(fit_sum)), function(i) {
  rlnorm(n_sim, meanlog = pars[i, "mu"], sdlog = pars[i, "sdlog"])
})

# Optional process layer: N_i | lambda_i ~ Poisson(lambda_i)
N_draws <- matrix(
  rpois(length(lambda_draws), lambda = as.vector(lambda_draws)),
  nrow = n_sim
)

sample_dogs_draw <- rowSums(N_draws)   # posterior draws of sampled total dogs

# 2) Keep your class composition logic
A_sample <- setNames(sampled_class_props$area_ha, sampled_class_props$class)
w_class  <- setNames(sampled_class_props$dog_share, sampled_class_props$class)

# 3) Build a deterministic multiplier per fokontany
# dogs_fkt_draw = sample_dogs_draw * k_fkt
k_fkt <- urban_str_fkt_all %>%
  transmute(
    fkt = fkt,
    area_ha = area / 10000,
    k = area_ha * (
      (w_class["green"]      / A_sample["green"])      * p_green +
      (w_class["industrial"] / A_sample["industrial"]) * p_industrial +
      (w_class["urban"]      / A_sample["urban"])      * p_urban +
      (w_class["water"]      / A_sample["water"])      * p_water
    )
  )

city_multiplier <- sum(k_fkt$k, na.rm = TRUE)

# 4) Posterior draws for city total
city_total_draw <- sample_dogs_draw * city_multiplier

# 5) Summary with uncertainty interval
city_summary <- c(
  mean   = mean(city_total_draw),
  median = median(city_total_draw),
  lcl95  = quantile(city_total_draw, 0.025),
  ucl95  = quantile(city_total_draw, 0.975)
)

city_summary
```

This goes too far maybe... uncertainty propagation from emmeans

```{r}
library(dplyr)
library(emmeans)
library(MASS)

set.seed(123)
n_sim <- 5000

# -----------------------------
# A) Draw sampled total dogs from model_6
# -----------------------------
fit_sum <- inla.nmix.lambda.fitted(model_6)$fitted.summary
# expected columns: mean.lambda, sd.lambda

mu_sd_to_lnorm <- function(m, s) {
  sigma2 <- log1p((s^2) / (m^2))
  mu <- log(m) - 0.5 * sigma2
  c(mu = mu, sdlog = sqrt(sigma2))
}

pars <- t(mapply(mu_sd_to_lnorm, fit_sum$mean.lambda, fit_sum$sd.lambda))

lambda_draws <- sapply(seq_len(nrow(fit_sum)), function(i) {
  rlnorm(n_sim, meanlog = pars[i, "mu"], sdlog = pars[i, "sdlog"])
})

# Optional process layer:
N_draws <- matrix(
  rpois(length(lambda_draws), lambda = as.vector(lambda_draws)),
  nrow = n_sim
)

sample_dogs_draw <- rowSums(N_draws)

# -----------------------------
# B) Draw class rates from emmeans uncertainty
# -----------------------------
# Replace glm_class with your fitted Poisson model used for emmeans
# Example: glm_class <- glm(count ~ class + offset(log(area_ha)), family = poisson, data = ...)
emm_link <- emmeans(m, ~ class, type = "link")

emm_df <- as.data.frame(emm_link)
mu_eta <- emm_df$emmean
Sigma_eta <- as.matrix(vcov(emm_link))

eta_draw <- MASS::mvrnorm(n = n_sim, mu = mu_eta, Sigma = Sigma_eta)
rate_draw <- exp(eta_draw)  # dogs per ha if your model was built as rate model with offset(log(area))

colnames(rate_draw) <- as.character(emm_df$class)

# -----------------------------
# C) Convert sampled class rates -> sampled dog shares each simulation
# -----------------------------
# Sample composition in the sampled zone
p_sample <- c(
  green = p_sample_green,
  industrial = p_sample_industrial,
  urban = p_sample_urban,
  water = p_sample_water
)

# Keep only required classes in a fixed order
rate_draw <- rate_draw[, c("green", "industrial", "urban", "water"), drop = FALSE]

# weights per simulation = p_k * rate_k
w_draw <- sweep(rate_draw, 2, p_sample, "*")
dog_share_draw <- w_draw / rowSums(w_draw)

# sampled class areas (ha)
A_sample <- c(
  green = sample_green,
  industrial = sample_industrial,
  urban = sample_urban,
  water = sample_water
)

# dogs per class in sampled zone (per simulation)
dogs_class_draw <- dog_share_draw * sample_dogs_draw

# class densities (dogs/ha) per simulation
dens_class_draw <- sweep(dogs_class_draw, 2, A_sample, "/")
# matrix n_sim x 4, columns = green industrial urban water

# -----------------------------
# D) Extrapolate to whole city for each simulation
# -----------------------------
city_df <- urban_str_fkt_all %>%
  transmute(
    fkt = fkt,
    area_ha = area / 10000,
    p_green = p_green,
    p_industrial = p_industrial,
    p_urban = p_urban,
    p_water = p_water
  )

# effective city area by class (ha)
city_class_area <- c(
  green = sum(city_df$area_ha * city_df$p_green, na.rm = TRUE),
  industrial = sum(city_df$area_ha * city_df$p_industrial, na.rm = TRUE),
  urban = sum(city_df$area_ha * city_df$p_urban, na.rm = TRUE),
  water = sum(city_df$area_ha * city_df$p_water, na.rm = TRUE)
)

# total dogs in city per simulation
# each sim: sum_k density_k(sim) * city_class_area_k
city_total_draw <- dens_class_draw %*% as.numeric(city_class_area)
city_total_draw <- as.numeric(city_total_draw)

city_summary <- c(
  mean = mean(city_total_draw),
  median = median(city_total_draw),
  lcl95 = quantile(city_total_draw, 0.025),
  ucl95 = quantile(city_total_draw, 0.975)
)

city_summary
```

### Extrapolation based on dog-to-human ratio

```{r}
sample_average_hum_dens <- sum(sitecovs_fkt$population) / sum(sitecovs_fkt$area_admin) * 10000
sample_humans <- sample_average_hum_dens * sum(sitecovs_fkt$area_real)/10000
sample_dogs / sample_humans

1350000*(sample_dogs / sample_humans) #human population in CUA in 2024
```

```{r}
set.seed(123)
n_sim <- 5000

# 1) Fixed human denominator in sampled area
sample_average_hum_dens <- sum(sitecovs_fkt$population) / sum(sitecovs_fkt$area_admin) * 10000
sample_humans <- sample_average_hum_dens * sum(sitecovs_fkt$area_real) / 10000

# 2) Draw sampled dogs from model_6 uncertainty
fit_sum <- inla.nmix.lambda.fitted(model_6)$fitted.summary
# expected columns: mean.lambda, sd.lambda

mu_sd_to_lnorm <- function(m, s) {
  sigma2 <- log1p((s^2) / (m^2))
  mu <- log(m) - 0.5 * sigma2
  c(mu = mu, sdlog = sqrt(sigma2))
}

pars <- t(mapply(mu_sd_to_lnorm, fit_sum$mean.lambda, fit_sum$sd.lambda))

lambda_draws <- sapply(seq_len(nrow(fit_sum)), function(i) {
  rlnorm(n_sim, meanlog = pars[i, "mu"], sdlog = pars[i, "sdlog"])
})

# Optional process variation: N_i | lambda_i ~ Poisson(lambda_i)
N_draws <- matrix(
  rpois(length(lambda_draws), lambda = as.vector(lambda_draws)),
  nrow = n_sim
)

sample_dogs_draw <- rowSums(N_draws)

# 3) Dog-to-human ratio draws and city extrapolation draws
ratio_draw <- sample_dogs_draw / sample_humans
city_dogs_draw <- 1350000 * ratio_draw

# 4) Summaries
ratio_summary <- c(
  mean = mean(ratio_draw),
  median = median(ratio_draw),
  lcl95 = quantile(ratio_draw, 0.025),
  ucl95 = quantile(ratio_draw, 0.975)
)

city_summary <- c(
  mean = mean(city_dogs_draw),
  median = median(city_dogs_draw),
  lcl95 = quantile(city_dogs_draw, 0.025),
  ucl95 = quantile(city_dogs_draw, 0.975)
)

ratio_summary
city_summary
```

# Additional considerations

### Do the poorly fitted fokontany have something in common?

```{r}
# Diagnose why some fokontany fit poorly in the observed-vs-predicted count graph.
# This separates mean mismatch from excessive within-site variability.

lambda_diag_df <- lambda_plot_df %>%
  transmute(
    site,
    lambda_mean = mean.lambda,
    lambda_lwr = quant025.lambda,
    lambda_upr = quant975.lambda
  )

site_diag_df <- obs_box_df %>%
  group_by(site) %>%
  summarise(
    obs_mean = mean(count),
    obs_median = median(count),
    obs_sd = sd(count),
    obs_min = min(count),
    obs_max = max(count),
    .groups = "drop"
  ) %>%
  left_join(pred_obs_df, by = "site") %>%
  left_join(lambda_diag_df, by = "site") %>%
  left_join(select(p_df, site, p_mean, p_lwr, p_upr), by = "site") %>%
  mutate(
    mean_gap = obs_mean - mean_lp,
    upper_gap = obs_max - upr_lp,
    lower_gap = lwr_lp - obs_min,
    pred_sd_approx = pmax((upr_lp - lwr_lp) / 3.92, 1e-6),
    spread_ratio = obs_sd / pred_sd_approx,
    likely_issue = case_when(
      obs_min > upr_lp ~ "All visits are higher than predicted: abundance likely underestimated or a site covariate is missing.",
      obs_max < lwr_lp ~ "All visits are lower than predicted: abundance likely overestimated or detection is too high.",
      abs(mean_gap) <= 0.5 & spread_ratio > 1.5 ~ "Average count is about right, but visit-to-visit variation is too large: missing detection heterogeneity or overdispersion.",
      mean_gap > 0 ~ "Observed counts are usually above prediction: lambda or p is too low.",
      mean_gap < 0 ~ "Observed counts are usually below prediction: lambda or p is too high.",
      TRUE ~ "No clear single cause."
    ),
    misfit_score = abs(mean_gap) + pmax(upper_gap, 0) + pmax(lower_gap, 0)
  ) %>%
  arrange(desc(misfit_score))

site_diag_df %>%
  select(site, obs_mean, obs_min, obs_max, mean_lp, lwr_lp, upr_lp, lambda_mean, p_mean, spread_ratio, likely_issue) %>%
  print(n = nrow(.))

worst_sites <- site_diag_df %>%
  slice_head(n = 6) %>%
  pull(site) %>%
  as.character()

visit_diag_df <- obs_plot_df %>%
  filter(as.character(site) %in% worst_sites) %>%
  left_join(select(pred_obs_df, site, mean_lp, lwr_lp, upr_lp), by = "site") %>%
  mutate(
    site = factor(site, levels = worst_sites),
    visit_residual = count - mean_lp
  )

ggplot(visit_diag_df, aes(x = visit, y = count, group = site)) +
  geom_hline(aes(yintercept = mean_lp), color = "#2e8b57", linewidth = 0.5) +
  geom_ribbon(aes(ymin = lwr_lp, ymax = upr_lp), fill = "#bfe3cf", alpha = 0.35) +
  geom_line(color = "#0b3c5d", linewidth = 0.45) +
  geom_point(color = "#0b3c5d", size = 1.6) +
  facet_wrap(~ site, scales = "free_y") +
  labs(
    title = "Visit-level counts for the most poorly fit fokontany",
    x = "Visit",
    y = "Observed count"
  ) +
  theme_minimal()
```

How to read this diagnostic:

-   If all five visits are mostly above the green band, the model is underestimating expected counts there. That usually points to missing abundance structure for that fokontany.
-   If the average count is close to the green line but one or two visits jump far away, the problem is more likely missing visit-level detection variation.
-   If the blue counts span a much wider range than the green band, the model is missing within-site overdispersion or temporal heterogeneity.
-   Compare `lambda_mean` and `p_mean`: a site with high observed counts but moderate `lambda_mean` and ordinary `p_mean` suggests underestimated abundance; a site with low observed counts but high `lambda_mean` may indicate overstated detection or abundance.

```{r}
# Do the poorly fit fokontany share similar covariate values?

bad_fit_sites <- site_diag_df %>%
  slice_head(n = 5) %>%
  transmute(site = as.character(site), misfit_score, likely_issue)

numeric_covars <- c(
  "length", "length_s", "pop_dens_s", "index1", "index2", "pcc_s", "pcc_total",
  "area_s", "area_real", "area_admin", "p_compact", "p_sparse", "p_ricefield",
  "p_industrial", "p_water", "percentage covared", "population"
)
numeric_covars <- intersect(numeric_covars, names(sitecovs_fkt))

covar_profile_df <- sitecovs_fkt %>%
  mutate(site = as.character(fkt)) %>%
  semi_join(select(site_diag_df, site), by = "site") %>%
  mutate(group = if_else(site %in% bad_fit_sites$site, "worst_fit", "other_sites")) %>%
  select(site, group, all_of(numeric_covars)) %>%
  pivot_longer(
    cols = all_of(numeric_covars),
    names_to = "covariate",
    values_to = "value"
  ) %>%
  mutate(value = suppressWarnings(as.numeric(value))) %>%
  filter(is.finite(value))

covar_group_summary <- covar_profile_df %>%
  group_by(covariate) %>%
  mutate(
    z_value = as.numeric(scale(value))
  ) %>%
  ungroup() %>%
  group_by(covariate, group) %>%
  summarise(
    n = n(),
    mean_value = mean(value),
    mean_z = mean(z_value),
    prop_above_global_median = mean(value > median(value, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = group,
    values_from = c(n, mean_value, mean_z, prop_above_global_median)
  ) %>%
  mutate(
    mean_diff = mean_value_worst_fit - mean_value_other_sites,
    z_gap = mean_z_worst_fit - mean_z_other_sites,
    median_direction_gap = prop_above_global_median_worst_fit - prop_above_global_median_other_sites
  ) %>%
  arrange(desc(abs(z_gap)))

covar_group_summary %>%
  select(covariate, n_worst_fit, n_other_sites, mean_value_worst_fit, mean_value_other_sites, z_gap, median_direction_gap) %>%
  print(n = nrow(.))

top_common_covars <- covar_group_summary %>%
  slice_head(n = 6) %>%
  pull(covariate)

bad_fit_heatmap_df <- covar_profile_df %>%
  filter(site %in% bad_fit_sites$site, covariate %in% top_common_covars) %>%
  group_by(covariate) %>%
  mutate(z_value = as.numeric(scale(value))) %>%
  ungroup() %>%
  mutate(
    site = factor(site, levels = bad_fit_sites$site),
    covariate = factor(covariate, levels = rev(top_common_covars))
  )

ggplot(bad_fit_heatmap_df, aes(x = site, y = covariate, fill = z_value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#b2182b", midpoint = 0) +
  labs(
    title = "Covariate profile of the worst-fit fokontany",
    x = "Fokontany",
    y = "Covariate",
    fill = "Within-covariate z-score"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
