---
title: "Data preparation - Dog population estimate in Tana"
author: "Katerina"
format: html
editor: visual
---

```         
```

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)

```

## Data import and adaptation

Import of the raw database downloaded from WVS server, creation of a new variable "fokontany" (sampled polygons), according to the team and dates worked

Note: two pairs of small neighboring fokontany were assessed as one polygon, this is the case of;

Faravohitra-Analakely = Faravohitra Mandrosoa and Soarano Ambondrona Ambodifilao

Isotry-Antetezana = Antetezanafovoany II and Cite Ambodin'Isotry

Thus we have 19 polygons /counting sites which represent 21 administrative units.

```{r}
com <- read.csv("DogCountDataRaw.csv")
summary(com)

com <- com %>%
    mutate(
    date = as.Date(`Date.Created`, format = "%d/%m/%Y"),
    fkt = case_when(
      date >= as.Date("2024-06-28") & date <= as.Date("2024-07-04") ~ "Andraisoro",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe01"  ~ "Mandroseza",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe02"  ~ "Mandroseza",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe03"  ~ "Mandroseza",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe04"  ~ "Ambohimiandra",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe05"  ~ "Ambohimiandra",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe06"  ~ "Ambohimiandra",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe07"  ~ "Ambohimiandra",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe08"  ~ "Ambohimiandra",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe09"  ~ "Ambohimiandra",
      date >= as.Date("2024-07-05") & date <= as.Date("2024-07-11") & UserName == "Equipe10"  ~ "Ambohimiandra",
      
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe02"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe03"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe04"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe05"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe06"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe07"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe08"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe09"  ~ "Ambodimita",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe01"  ~ "Ambohimanandray",
      date >= as.Date("2024-07-12") & date <= as.Date("2024-07-18") & UserName == "Equipe10"  ~ "Ambohimanandray",
      
      date >= as.Date("2024-07-19") & date <= as.Date("2024-07-25")  ~ "Analamahitsy",
      
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe05"  ~ "Andraharo",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe07"  ~ "Andraharo",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe08"  ~ "Andraharo",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe02"  ~ "Ambodivona",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe04"  ~ "Ambodivona",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe01"  ~ "Anjanahary",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe03"  ~ "Anjanahary",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe06"  ~ "Anjanahary",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe09"  ~ "Anjanahary",
      date >= as.Date("2024-07-26") & date <= as.Date("2024-08-01") & UserName == "Equipe10"  ~ "Anjanahary",
      
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe05" ~ "Isotry-Antetezana",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe07" ~ "Isotry-Antetezana",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe02"  ~ "Faravohitra-Analakely",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe03"  ~ "Faravohitra-Analakely",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe04"  ~ "Faravohitra-Analakely",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe01"  ~ "Faravohitra-Analakely",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe06"  ~ "Mahamasina",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe08"  ~ "Mahamasina",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe09"  ~ "Mahamasina",
      date >= as.Date("2024-08-02") & date <= as.Date("2024-08-08") & UserName == "Equipe10"  ~ "Mahamasina",
      
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe06"  ~ "Andavamamba",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe05"  ~ "Andavamamba",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe07"  ~ "Andavamamba",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe04"  ~ "Anosipatrana",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe09"  ~ "Anosipatrana",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe10"  ~ "Anosipatrana",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe01"  ~ "Ankazatoho",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe02"  ~ "Ankazatoho",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe03"  ~ "Ankazatoho",
      date >= as.Date("2024-08-09") & date <= as.Date("2024-08-16") & UserName == "Equipe08"  ~ "Ankazatoho",

      date >= as.Date("2024-08-19") & date <= as.Date("2024-08-23")  ~ "Ankerana",
      
      date >= as.Date("2024-08-26") & date <= as.Date("2024-08-30")  ~ "Soavimasoandro",
      
      date >= as.Date("2024-10-10") & date <= as.Date("2024-10-16") ~ "Ampandrana",
      date >= as.Date("2024-10-17") & date <= as.Date("2024-10-31") ~ "Ambatomaro",
                  TRUE ~ NA
    )
  )
View(com)
```

Quality check of the previous step (dates with no fokontany attributed will represent dates of trial work, not to be included downstream)

```{r}
#check if all fkt exist
unique(com$fkt)

#check if all days have been asigned to a fkt
com %>% filter(is.na(fkt)) %>% distinct(date)
```

## Summarizing and visualizing the count data

### Daily summaries per fokontany

```{r}
#retain only variables pertaining to count, fokontany and date
com_clean <- com %>% 
  mutate (n = nombre.de.chien) %>%
  dplyr::select(date, n, fkt) %>% 
  filter(!is.na(fkt))

#sum per day per fokontany
sumfktday <-com_clean %>%
  group_by(date, fkt) %>%
  summarise(sum_fkt = sum(n, na.rm = TRUE))

#mean, min, max and variance of the daily counts per fokontany
summary_fkt <-sumfktday %>%
  group_by(fkt) %>%
  summarise(minfkt= min(sum_fkt),
            maxfkt= max(sum_fkt),
            meanfkt = mean(sum_fkt),
            varfkt = var(sum_fkt) 
            ) %>%
  mutate(id = c(1:19))

#visualize
summary_fkt

boxplot(sum_fkt ~ fkt, data = sumfktday, las = 2, cex.axis = 0.5)

```

### Daily summaries per path (team)

```{r}
#retain only count, fokontany and date
com_clean_transect <- com %>% 
  mutate (
    n = nombre.de.chien,
    transect = paste(fkt, UserName), 
    date = as.Date(date, format = "%d/%m/%Y")) %>%
  select(date, n, fkt, transect) %>% 
  arrange(date, transect) %>%
  filter(!is.na(fkt)) 

head(com_clean_transect)
unique(com_clean_transect$transect)

#sum per day per transect
sum_transect_day <-com_clean_transect %>%
  group_by(fkt, transect, date) %>%
  summarise(sum_transect = sum(n, na.rm = TRUE))

#mean, min, max and variance of the daily counts per transect
summary_transect <- sum_transect_day %>%
  group_by(transect) %>%
  summarise(
    mean_transect = mean(sum_transect, na.rm = TRUE),
    max_transect = max(sum_transect, na.rm = TRUE),
    min_transect = min(sum_transect, na.rm = TRUE),
    variance_transect = var(sum_transect, na.rm = TRUE)
  )

summary_transect

```

## Count matrices

### Matrix of counts using fokontany as site

```{r}
sfd_ranks <- sumfktday %>%
  group_by(fkt) %>%
  arrange(fkt, date) %>%
  mutate(repetition = row_number())

obs_matrix_fkt <- sfd_ranks %>%
  dplyr::select(fkt, sum_fkt, repetition) %>%
  pivot_wider(
    names_from = repetition,
    values_from = sum_fkt) %>%
  ungroup() %>% 
  dplyr::select(- fkt) %>%
  as.matrix()
  
obs_matrix_fkt

write.csv(obs_matrix_fkt, "obs_matrix_fkt.csv")
```

### Matrix of counts using path as site

```{r}
sum_transect_day_ranked <- sum_transect_day %>%
  group_by(transect) %>%
  mutate(repetition = row_number()) %>%
  ungroup()

obs_matrix_transect <- sum_transect_day_ranked %>%
  select(transect, repetition, sum_transect) %>%
  pivot_wider(
    names_from = repetition,
    values_from = sum_transect) %>%
  ungroup() %>% 
  select(- transect) %>%
  as.matrix()
 
obs_matrix_transect

write.csv(obs_matrix_transect, "obs_matrix_transect.csv")
```
